---
title: "Exploring Patterns and Predictors of Infant Mortality Across the Globe"
subtitle: "DANL-320 Project Write-up"
author: Nada Trabelsi
date: 2026-05-14

toc: true
toc-title: "Table of Contents"
toc-depth: 2
code-fold: true
fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false
---

```{r setup}
#| include: false

# Loading pkgs
library(readr)
library(tidyverse)
library(readxl)
library(skimr)
library(dplyr)
library(sf)
library(ggthemes)
library(ggtext)
library(ggrepel)
library(plotly)
library(knitr)
library(zoo)
#install.packages("fixest")
library(fixest)
library(broom)
library(glmnet)
library(dplyr)
library(factoextra)
library(cluster)
library(fpc)
library(plm)
library(rmarkdown)



# set theme

#main data set
global_health <- read_csv("global_health.csv") 

#2nd data set
birth_rate_df <- read_excel("P_Data_Extract_From_World_Development_Indicators_all_years_for_kaggle.xlsx")  |> 
  mutate(Year = Time,
         Country = Country_name) |> 
  select(Country, Year, Birth_rate)

#final cleaned dataset
global_health <- global_health |> 
  left_join(birth_rate_df) |> 
  mutate(Infant_Mortality_Rate = round((Infant_Deaths / (Birth_rate / 1000 * Total_Population)) * 1000, 2)) |> 
  select(-Water_Access_Percent, -Suicide_Rate_Percent, -Alcohol_Consumption_Per_Capita, -Hospital_Beds_Per_1000, -Obesity_Rate_Percent, -Overweight_Rate_Percent, -Underweight_Rate_Percent, -Tuberculosis_Per_100000, -Life_Expectancy_Female, -Life_Expectancy_Male, -CO2_Exposure_Percent, -Labour_Force_Total, -Unemployment_Rate) |> 
  arrange(Country, Year) |> 
  group_by(Country) |> 
  mutate(across(where(is.numeric), ~ na.locf(., na.rm = FALSE))) |> 
  drop_na() |> 
  ungroup()

# split to train and testing 
global_healthsplit <- global_health |> 
  mutate(rnd = runif(n()))

dtrain <- global_healthsplit |> 
  filter(rnd >= .3)

dtest <- global_healthsplit |> 
  filter(rnd < .3)

# Regularized regression

## following code generated by ChatGPT to fix issue in predicting regularized models later on 

x <- model.matrix(
  Infant_Mortality_Rate ~ 
    Immunization_Rate +
    log(GDP_Per_Capita) +
    Fertility_Rate +
    Air_Pollution +
    Safe_Water_Access_Percent +
    Sanitary_Expense_Per_GDP +
    Urban_Population_Percent + 
    factor(Country) +
    factor(Year),
  data = dtrain
)[, -1]

x_test <-model.matrix(
  Infant_Mortality_Rate ~ 
    Immunization_Rate +
    log(GDP_Per_Capita) +
    Fertility_Rate +
    Air_Pollution +
    Safe_Water_Access_Percent +
    Sanitary_Expense_Per_GDP +
    Urban_Population_Percent + 
    factor(Country) +
    factor(Year),
  data = dtest
)[, -1]

y <- dtrain$Infant_Mortality_Rate - 1

y_test <-dtest$Infant_Mortality_Rate - 1

## following code generated by ChatGPT to fix issue in predicting regularized models later on 


missing_cols <- setdiff(colnames(x), colnames(x_test))

for(col in missing_cols) {
  x_test <- cbind(x_test, 0)
  colnames(x_test)[ncol(x_test)] <- col
}

# remove extra columns
extra_cols <- setdiff(colnames(x_test), colnames(x))

if(length(extra_cols) > 0) {
  x_test <- x_test[, !(colnames(x_test) %in% extra_cols), drop = FALSE]
}

# reorder to EXACTLY match training matrix
x_test <- x_test[, colnames(x), drop = FALSE]


``` 

# Introduction
Infant mortality is the death of a child under 1 year old, measured in deaths per 1,000 live births. For example, a rate of 9 indicates that 9 infants die for every 1,000 live births, not accounting for stillbirths or miscarriages. Infant death can be caused by several different reasons and can be caused by accidents, lack of resources, and severe living conditions (World Health Organization). 

The infant mortality rate (IMR) is an important health indicator as it is a marker for a country's overall public health, development, and social equity. Factors such as poor access to clean water, low immunization coverage, and limited economic development can have a major impact on a country's ability to reduce infant mortality. This project aims to analyze the association between IMR and different factors, and assess their ability to predict IMR. 

# Data

## Source and Scope

The main data set, "Global Health and Development (2012-2021)", was produced by Martina Galasso by combining reliable data from the World Bank and the World Health Organization (WHO) to showcase a comprehensive overview of of global health, demographics, economic, and environmental metrics for 188 countries between 2012 and 2021. This data set included the total number of infant deaths in a year but did not include the infant mortality rate. This led me to search for another data set that provided an indicator that would allow me to calculate the infant mortality rate. The second data set, "World Bank World Development Indicators", curated by Nicolás Ariel González Muñoz contained world development indicators from 1960 to 2022. This data set contained the birth rate, which allowed me to reverse the equations and find the infant mortality rate for every available country in the latest available year in the first data set. Both data sets were combined, resulting in a final panel data spanning 2012-2021, with each row representing one country at a specific year.


## Cleaning and Preparation
After joining both datasets, `Infant_Mortality_Rate` was calculated by dividing total infant deaths with total live births and multiplying by 1,000. To find the total number of live births, the Birth_Rate was divided by 1,000 and multiplied by the Total_Population. The infant mortality rate was then found by dividing the total live births with the Infant_Deaths and multiplying by 1,000. The rate was rounded to two decimal points to keep the data set from looking cluttered. 

Variables with high levels of missing values were removed and remaining observations with missing values were originally estimated using the last observation carried forward with the `na.locf` function. The dataset was also later standardized with the `scaled()` for clustering. 

## Variables
This study focuses on the following factors as predictors of IMR:

- `GDP_Per_Capita`: The total value of a country's economic output per person 
- `Sanitary_Expense_per_GDP`: The portion of GDP spent on sanitary services, e.g., waste management
- `Ferility_Rate`: Birth rate per woman
- `Immunization_Rate`: The percentage of the population who received vaccinations to protect against certain diseases
- `Safe_Water_Access`: The percentage of the population who have access to clean and safe water 
- `Urban_Population_percent`: The percentage of the population that live in urban areas


## Exploratory Analysis
```{r}
#| echo: false
global_health |> 
  select(Infant_Mortality_Rate,
         GDP_Per_Capita,
         Sanitary_Expense_Per_GDP,
         Fertility_Rate,
         Immunization_Rate,
         Safe_Water_Access_Percent,
         Urban_Population_Percent
         ) |> 
  skim()
```

The average IMR is about 22 around but there is great variation between the third quarter and the maximum IMR, jumping from 34 to 86. There seems to be great variation and range in key variables between countries.

```{=html}
<div style="text-align: center; margin: 1em 0;">
  <iframe
    src="https://0196e215-b119-0a0b-0be6-179545105a82.share.connect.posit.cloud/"
    style="width: 1100px; height: 800px; border: none;"
  ></iframe>
</div>
```

A majority of countries that experience severe infant mortality are Central and Western Africa and South Asia. 

```{r}
#| echo: false
gh3 <- global_health |> 
ggplot(gh3,
       mapping = aes(x = Fertility_Rate,
                     y = Infant_Mortality_Rate)) +
  geom_smooth(method = "loess", se = F, linewidth = 1) +
  geom_point(aes(color = Safe_Water_Access_Percent	)) +
  labs(title = "Fertility Rate vs. IMR",
       caption = "Source: World Bank")+
  xlab("Ferility Rate")+
  ylab("Infant Mortality Rate") +
  scale_x_continuous(labels = scales::percent_format(scale = 1))+
  theme_bw()
gh3
```
There appears to be a strong, positive relation between fertility rate and IMR. Countries with higher access to safe water typically have lower IMR, while countries with low access to safe water higher IMR.

# Supervised Machine Learning Analysis
The dataset was split into train and test datasets in order to evaluate how well the models are able to predict IMR outside of the training data. Root mean squared error (RMSE) and Mean squared error (MAE) will be used to evaluate the models performance to predict IMR. Several regression models are utilized to examine the relationship between selected predictors and IMR. 

## Two-Way Fixed Effects Model

A two-way fixed effects model is used because it control for unobserved differences across countries while controlling for yearly trends and/or shocks. 

$$
Infant_Mortality_Rate_{it} &= \beta_0 + \textbf{X'_{it}\beta} + \alpha_i + \lambda_t + \varepsilon_{it},
$$

where the outcome is IMR in country $i$ at time $t$, $\alpha_i$ captures country fixed effects, $\lambda_t$ captures year fixed effects, and $\varepsilon_{it}$ is the error term.

```{r}
#| echo: false

twfe <- pdata.frame(dtrain, index = c("Country", "Year"))
model_plm <- plm(
  Infant_Mortality_Rate ~ 
    Immunization_Rate +
    log(GDP_Per_Capita) +
    Fertility_Rate +
    Air_Pollution +
    Safe_Water_Access_Percent +
    Sanitary_Expense_Per_GDP+
    Urban_Population_Percent,
  
  data = twfe,
  model = "within",
  effect = "twoways"
)

m1_coef <- tidy(model_plm) |> 
  filter(!grepl("Country|Year", term))

ggplot(m1_coef, aes(x = estimate, y = reorder(term, estimate))) +
  geom_point(color = "blue") +
  geom_errorbarh(
    aes(xmin = estimate - 1.96 * std.error,
        xmax = estimate + 1.96 * std.error),
    height = 0.2
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(,
    x = "Coefficient Estimate",
    y = ""
  ) +
  theme_minimal()


```

Fertility rate has the strongest association among all predictors, iplying increasing fertility by one, increases the IMR by about 6.5 deaths, holding all other factors constant. GDP per capita is the predictor with the second highest effect, with a coefficient of about 1.2. This means thatthat a 10% increase in GDP per capita is associatied with a 0.12 increase in IMR, holding everything else constant. All the other predictors are near zero indicating they have very weak association with IMR after controlling for the other predictors. 

## Lasso Linear Regression
Lasso regression is used to address any multicollinearity as it performs strict variable selection, shrinking less important predictors coefficients toward zero. Country and year dummy variables are used as fixed effects to control for unobserved differences between countries and year trends and/or shocks. 

```{r}
#| echo: false
lasso <- cv.glmnet(x, y, alpha = 1)

coef_df <- as.matrix(coef(lasso, s = "lambda.min")) %>%
  as.data.frame()

coef_df$variable <- rownames(coef_df)
colnames(coef_df)[1] <- "coefficient"

coef_df <- coef_df %>%
  filter(variable != "(Intercept)",
         !grepl("factor", variable)) %>%
  arrange(coefficient)

ggplot(coef_df, aes(x = coefficient, y = reorder(variable, coefficient))) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Lasso Regression Coefficients",
    x = " ",
    y = " "
  ) +
  theme_minimal()
```

Similar to the two-way fixed effects model, fertility rate has the strongest, positive coefficient, meaning increasing fertility by one, increases the IMR by about 7.8 deaths, holding all other factors constant. The lasso regression has similar results to the first model, however, log(GDP_Per_Capita)'s coefficients are reversed indicating that a 10% increase in GDP per capita is associatied with a 0.1 decrease in IMR, holding everything else constant. All other variables are near zero indicating they are very weak predictors of IMR. 

## Ridge Linear Regression
Ridge regression is used to address any multicollinearity by shrinking coefficients, similar to Lasso, but instead stabilizing the shared effect among the highly correlated variables. 
```{r}
#| echo: false
ridge <- cv.glmnet(x, y, alpha = 0)

coef_df <- as.matrix(coef(ridge, s = "lambda.min")) %>%
  as.data.frame()

coef_df$variable <- rownames(coef_df)
colnames(coef_df)[1] <- "coefficient"

coef_df <- coef_df %>%
  filter(variable != "(Intercept)",
         !grepl("factor", variable)) %>%
  arrange(coefficient)

ggplot(coef_df, aes(x = coefficient, y = reorder(variable, coefficient))) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Lasso Regression Coefficients",
    x = " ",
    y = " "
  ) +
  theme_minimal()
```
Compared to the lasso model, the ridge model shrinks the effect of the fertility rate to about 3.5, meaning increasing fertility by one, increases the IMR by about 3.5 deaths, holding all other factors constant. The effect of the GDP per capita is increased, showing a stronger negative effect on IMR compared to the lasso model. Sanitary expense per GDP is shrunk completely to zero, showing it does not contribute meaningfully when the other factors are considered. 

## Comparing the Models
```{r}
#| echo: false
pred_plm_test <- predict(model_plm, newdata = dtest)
pred_lasso_test <- as.numeric(predict(lasso, newx = x_test, s = "lambda.min"))
pred_ridge_test <- as.numeric(predict(ridge, newx = x_test, s = "lambda.min"))

model_performance <- tibble(
  model = c(
    "2-Way FE regression",
    "Lasso regression",
    "Ridge regression"
  ),
  rmse = c(
    sqrt(mean((y_test - pred_plm_test)^2)),
    sqrt(mean((y_test - pred_lasso_test)^2)),
    sqrt(mean((y_test - pred_ridge_test)^2))
  ),
  mae = c(
    mean(abs(y_test - pred_plm_test)),
    mean(abs(y_test - pred_lasso_test)),
    mean(abs(y_test - pred_ridge_test))
  )
) |> 
  arrange(rmse)

model_performance |> 
  mutate(
    rmse = round(rmse, 4),
    mae = round(mae, 4)
  ) |> 
  paged_table()
```
Lasso had the lowest prediciton error, followed closely by ridge regression. This is because the regularized regression model were able to shrink coefficients, which improves the models performance. The linear regression had the highest error, meaning compared to regularized regression it is not well at generalizing to fit data outside of the training data. 


# Unsupervised Machine Learning Analysis

K-means clustering will be used to group countries together in 4 distinct groups based on similar characterists to create cluster profiles. This will be used to identify distinct patterns among countries with low and high IMR. All observations are standardized because k-means clustering uses distance and variables need to be scaled so variables with larger scales do not dominate. 

```{r}
#| echo: false
data_cluster <- global_health |> 
  select(Infant_Mortality_Rate, Fertility_Rate, Air_Pollution, Sanitary_Expense_Per_GDP, Urban_Population_Percent, Immunization_Rate, Safe_Water_Access_Percent, GDP_Per_Capita)

cluster_scaled <- data_cluster |> 
  scale()

km4 <- kmeans(cluster_scaled, centers = 4, nstart = 25)

data_cluster$cluster <- km4$cluster

cluster_summary <- data_cluster |> 
  group_by(cluster) |> 
  summarise(across(everything(), ~round(mean(.x, na.rm = TRUE), 2)))

cluster_summary<- cluster_summary |> 
  pivot_longer(-cluster, names_to = "Predictor", values_to = "value") |> 
  pivot_wider(names_from = cluster, values_from = value) 

cluster_summary
```









